---
title: "Homework 1"
author: "Thiago de Araujo - UNI tbd2117"
output: github_document
---

```{r message=FALSE}
library(tidyverse)
```

### Problem 1

Here's the first code chunk to create the requested **dataframe** and its **means**:

```{r chunk1}
prob1 = 
  tibble(
    samp = rnorm(10),
    logic = samp > 0,
    char = c("This", "is", "a", "vector", "that", "contains", "a", "length", "of", "ten"),
    factor = factor(c("low", "medium", "high", "medium", "medium", "low", "medium", "low", "high", "low"))
)

# attempt to take the mean of each data variable
mean(pull(prob1, samp))
mean(pull(prob1, logic))
mean(pull(prob1, char))
mean(pull(prob1, factor))
```

Taking the mean of each variable works for the _samp_ variable (mean = `r mean(pull(prob1, samp))`) and for the _logic_ variable (mean = `r mean(pull(prob1, logic))`). It does not work for the char and factor variables.

```{r chunk1.5, eval=FALSE}
# appling the as.numeric function
as.numeric(pull(prob1, logic))
as.numeric(pull(prob1, char))
as.numeric(pull(prob1, factor))
````
Using the _as.numeric_ function it is possible to see values being attributed to TRUE and FALSE in the _logic_ variable and for each level in the _factor_ variable. It does help explain what happens when trying to take the mean of logical vectors (means of 0s and 1s).

Here's the **second code chunk**:
```{r chunk2}
as.numeric(pull(prob1, logic)) * pull(prob1, samp)
as.numeric(pull(prob1, char)) * pull(prob1, samp)
as.numeric(pull(prob1, factor)) * pull(prob1, samp)
```

### Problem 2

```{r}
# loading the penguins dataset
data("penguins", package = "palmerpenguins")
levels(penguins$species)
```

The _penguins_ dataset contains data on penguin species (`r levels(penguins$species)`), islands (`r levels(penguins$island)`), length of bills and flippers, body mass, sex, and year.

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. 

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm=TRUE)`.

```{r message=FALSE}
# scatterplot of flipper vs. bill length
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point()

# exporting scatterplot 
ggsave("scatter_plot.pdf", height = 4, width = 6)
```